---
title: "Assignment 2"
author: "munan"
date: "September 16, 2016"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


## Exercise 0 Basic Arithmetic


1. Assignment, variables
Try some simple commands

```{r e0q1}
a <- 2
b <- 50
c <- 3

a+b
```

```{r}
b/a
```

```{r}
50%%3
```

```{r}
sin(0)
```

```{r}
cos(0)
```

```{r}
sin45 <- sin(pi/4)
4*asin(sin45)
```

```{r}
e3 <- exp(3)
log(e3)
```

-     Minus, can be unary or binary
+     Plus, can be unary or binary
!     Unary not
~     Tilde, used for model formulae, can be either unary or binary
?     Help
:     Sequence, binary (in model formulae: interaction)
*     Multiplication, binary
/     Division, binary
^     Exponentiation, binary
%x%   Special binary operators, x can be replaced by any valid name
%%    Modulus, binary
%/%   Integer divide, binary
%*%   Matrix product, binary
%o%   Outer product, binary
%x%   Kronecker product, binary
%in%  Matching operator, binary (in model formulae: nesting)
<     Less than, binary
>     Greater than, binary
==    Equal to, binary
>=    Greater than or equal to, binary
<=    Less than or equal to, binary
&     And, binary, vectorized
&&    And, binary, not vectorized
|     Or, binary, vectorized
||    Or, binary, not vectorized
<-    Left assignment, binary
->    Right assignment, binary
$     List subset, binary
  

2. Try some basic Vector operations

```{r e0q2}
# create 
a_vec <- c(3, 50, 5, 6, 10, 40, 20, 2, 100, 30, 25, 80) 
# length 
length(a_vec)
```

```{r}
# sum
sum(a_vec)
```

```{r}
# cumulative sum
cumsum(a_vec)
```

```{r}
# summary statistics
summary(a_vec)
```

```{r}
# variance
var(a_vec)
```

```{r}
# is it reatlly a vector?
str(a_vec)
```

```{r}
# was that not clear?
is.vector(a_vec)
```

```{r}
# assign the 5th element of a_vec to d
d <- a_vec[5]
a_vec
```

```{r}
# what does this mean?
length(d)
```

```{r}
# Really? Well, ...
is.vector(d)
```

```{r}
# how about random numbers?
random_vec <- runif(20)
length(random_vec)
```

```{r}
random_vec
```

```{r}
# what if random_vec was really big?
random_vec_big <- runif(10000)
length(random_vec_big)
```

```{r}
head(random_vec_big)
```

```{r}
# need more than 6?
head(random_vec_big, 20)
```

```{r}
# too many decimal places?
options(digits = 2)
tail(random_vec_big)
```

```{r}
# What if I want my random series to be repeatable?
set.seed(2)
r <- runif(30)
head(r,10)
```

```{r}
r1 <- runif(20)
head(r1,10)
```

```{r}
set.seed(2)
s <- runif(10)
s
```

```{r}
# not enough digits?
options(digits = 4)
head(r)
```

```{r}
head(s)
```

```{r}
# can I get a picture of that?
plot(r)
```

```{r}
# that's not very helpful. Histogram.
hist(r)
```

```{r}
# doesn't look uniform? increase the sample size
hist(random_vec_big)
```

All these blocks of R may not be what you need.
Remember a_vec? Sure, avec is (3, 50, 5, 6, 10, 40, 20, 2, 100, 30, 25, 80). Here’s a_vec’s mean 30.9167.
Wow – a_vec is a strange set of data, what does it look like?

```{r}
boxplot(a_vec)
```


3. Finally, let’s look at matrice, lists, and data.frames

```{r}
# make a random matrix 
A <- matrix(data = floor(30*runif(16)), nrow = 4) 
A
```

```{r}
# how big is the matrix?
dim(A)
```

```{r}
# Let's make a vector to multiply by A
b <- floor(20*runif(dim(A)[2]))
b
```

```{r}
# solve Ax = b
x <- solve(A,b)
x
```

```{r}
b1 <- A%*%x
b1
```

```{r}
b2 <- as.vector(A%*%x) #coerce b2 into a vector (column convert to roll)
b2
```

```{r}
# inner product dot(e,f) = e'f
e <- floor(20*runif(4))
f <- floor(20*runif(4))
e
```

```{r}
f
```

```{r}
crossprod(e,f)
```

```{r}
sum(e*f)
```

```{r}
t(e)%*%f #transfer e crossprod f?
```

```{r}
# outer product
outer(e,f,"*")
```

```{r}
outer(e,f,">")
```

```{r}
#using automatic coersion
outer(e,f,">")*1
```



## Exercise 1 Vectors


1. Create the vectors: 

(a) (1,2,3,...,19,20) 
```{r e1q1}

```
(b) (20,19,...,2,1) 
(c) (1,2,3,...,19,20,19,18,...,2,1) 
(d) (4,6,3) and assign it to the name tmp. 
For parts (e), (f) and (g) look at the help for the function rep. 
(e) (4,6,3, 4,6,3,...,4,6,3) where there are 10 occurrences of 4. 
(f) (4,6,3, 4,6,3,...,4,6,3,4) where there are 11 occurrences of 4, 10 occurrences of 6 and 10 occurrences of 3. 
(g) (4,4,...,4, 6,6,...,6, 3,3,...,3) where there are 10 occurrences of 4, 20 occurrences of 6 and 30 occurrences of 3.


