---
title: "Assignment 2"
author: "Munan Hou"
date: "September 16, 2016"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


## Exercise 0 Basic Arithmetic


1. Assignment, variables
Try some simple commands

```{r e0q1}
a <- 2
b <- 50
c <- 3

a+b
```

```{r}
b/a
```

```{r}
50%%3
```

```{r}
sin(0)
```

```{r}
cos(0)
```

```{r}
sin45 <- sin(pi/4)
4*asin(sin45)
```

```{r}
e3 <- exp(3)
log(e3)
```

-     Minus, can be unary or binary
+     Plus, can be unary or binary
!     Unary not
~     Tilde, used for model formulae, can be either unary or binary
?     Help
:     Sequence, binary (in model formulae: interaction)
*     Multiplication, binary
/     Division, binary
^     Exponentiation, binary
%x%   Special binary operators, x can be replaced by any valid name
%%    Modulus, binary
%/%   Integer divide, binary
%*%   Matrix product, binary
%o%   Outer product, binary
%x%   Kronecker product, binary
%in%  Matching operator, binary (in model formulae: nesting)
<     Less than, binary
>     Greater than, binary
==    Equal to, binary
>=    Greater than or equal to, binary
<=    Less than or equal to, binary
&     And, binary, vectorized
&&    And, binary, not vectorized
|     Or, binary, vectorized
||    Or, binary, not vectorized
<-    Left assignment, binary
->    Right assignment, binary
$     List subset, binary
  

2. Try some basic Vector operations

```{r e0q2}
# create 
a_vec <- c(3, 50, 5, 6, 10, 40, 20, 2, 100, 30, 25, 80) 
# length 
length(a_vec)
```

```{r}
# sum
sum(a_vec)
```

```{r}
# cumulative sum
cumsum(a_vec)
```

```{r}
# summary statistics
summary(a_vec)
```

```{r}
# variance
var(a_vec)
```

```{r}
# is it reatlly a vector?
str(a_vec)
```

```{r}
# was that not clear?
is.vector(a_vec)
```

```{r}
# assign the 5th element of a_vec to d
d <- a_vec[5]
a_vec
```

```{r}
# what does this mean?
length(d)
```

```{r}
# Really? Well, ...
is.vector(d)
```

```{r}
# how about random numbers?
random_vec <- runif(20)
length(random_vec)
```

```{r}
random_vec
```

```{r}
# what if random_vec was really big?
random_vec_big <- runif(10000)
length(random_vec_big)
```

```{r}
head(random_vec_big)
```

```{r}
# need more than 6?
head(random_vec_big, 20)
```

```{r}
# too many decimal places?
options(digits = 2)
tail(random_vec_big)
```

```{r}
# What if I want my random series to be repeatable?
set.seed(2)
r <- runif(30)
head(r,10)
```

```{r}
r1 <- runif(20)
head(r1,10)
```

```{r}
set.seed(2)
s <- runif(10)
s
```

```{r}
# not enough digits?
options(digits = 4)
head(r)
```

```{r}
head(s)
```

```{r}
# can I get a picture of that?
plot(r)
```

```{r}
# that's not very helpful. Histogram.
hist(r)
```

```{r}
# doesn't look uniform? increase the sample size
hist(random_vec_big)
```

All these blocks of R may not be what you need.
Remember a_vec? Sure, avec is (3, 50, 5, 6, 10, 40, 20, 2, 100, 30, 25, 80). Here’s a_vec’s mean 30.9167.
Wow – a_vec is a strange set of data, what does it look like?

```{r}
boxplot(a_vec)
```


3. Finally, let’s look at matrice, lists, and data.frames

```{r}
# make a random matrix 
A <- matrix(data = floor(30*runif(16)), nrow = 4) 
A
```

```{r}
# how big is the matrix?
dim(A)
```

```{r}
# Let's make a vector to multiply by A
b <- floor(20*runif(dim(A)[2]))
b
```

```{r}
# solve Ax = b
x <- solve(A,b)
x
```

```{r}
b1 <- A%*%x
b1
```

```{r}
b2 <- as.vector(A%*%x) #coerce b2 into a vector (column convert to roll)
b2
```

```{r}
# inner product dot(e,f) = e'f
e <- floor(20*runif(4))
f <- floor(20*runif(4))
e
```

```{r}
f
```

```{r}
crossprod(e,f)
```

```{r}
sum(e*f)
```

```{r}
t(e)%*%f #transfer e crossprod f?
```

```{r}
# outer product
outer(e,f,"*")
```

```{r}
outer(e,f,">")
```

```{r}
#using automatic coersion
outer(e,f,">")*1
```



## Exercise 1 Vectors


1. Create the vectors: 

(a) (1,2,3,...,19,20) 

```{r e1q1}
e1q1a <- c(1:20)
```

(b) (20,19,...,2,1)

```{r}
e1q1b <- c(20:1)
```

(c) (1,2,3,...,19,20,19,18,...,2,1) 

```{r}
e1q1c <- c(c(1:20),c(19:1))
head(e1q1c,39)
```

(d) (4,6,3) and assign it to the name tmp. 

```{r}
tmp <- c(4,6,3)
```

For parts (e), (f) and (g) look at the help for the function rep. 

(e) (4,6,3, 4,6,3,...,4,6,3) where there are 10 occurrences of 4. 

```{r}
e1q1e <- c(rep(tmp, time=10))
e1q1e
```

(f) (4,6,3, 4,6,3,...,4,6,3,4) where there are 11 occurrences of 4, 10 occurrences of 6 and 10 occurrences of 3. 

```{r}
e1q1f <- c(rep(tmp, length=31))
e1q1f
```

(g) (4,4,...,4, 6,6,...,6, 3,3,...,3) where there are 10 occurrences of 4, 20 occurrences of 6 and 30 occurrences of 3.

```{r}
e1q1g <- c(rep(tmp, c(10,20,30)))
e1q1g
```


2. Create a vector of the values of e^x*cos(x) at x = 3,3.1,3.2,...,6. Make a plot of the multiplicants and the product.

```{r}
par(new=TRUE)
x <- seq(3,6,by=0.1)
prod <- exp(x)*cos(x)
plot(prod, col="blue", type="l") #(type set to "line")
points(cos(x),col="red")
points(exp(x),col="green")
```


3. Create the following vectors:

(a) (0.1^3*0.2^1, 0.1^6*0.2^4, ... , 0.1^36*0.2^34) 

```{r}
x <- seq(3, 36, by=3)
y <- seq(1, 34, by=3)
a <- (0.1^x)*(0.2^y)
a
```

(b) (2, (2^2)/2, (2^3)/3, ... , (2^25)/25)

```{r}
x <- seq(1:25)
b <- (2^x)/x
b
```


4. Calculate the following

(a) summary of (i^3+4*i^2), where i ranges from 10 to 100.

```{r}
i <- c(10:100)
a <- sum(i^3+4*i^2)
a
```

(b) summary of ((2^i)/i + (3^i)/(i^2)), where i ranges from 1 to 25.

```{r}
i <- (1:25)
b <- sum((2^i)/i + (3^i)/(i^2))
b
```


5. Use the function "paste"" to create the following character vectors of length 30: 

(a) ("label 1", "label 2", ....., "label 30"). Note that there is a single space between label and the number following. 

```{r}
paste("label", 1:30)
```

(b) ("fn1", "fn2", ..., "fn30"). In this case, there is no space between fn and the number following. 

```{r}
paste0("fn", 1:30)
```


6. Execute the following lines which create two vectors of random integers which are chosen with replacement from the integers 0, 1, ..., 999. Both vectors have length 250. 

```{r}
set.seed(50)
xVec <- sample(0:999, 250, replace=T)
yVec <- sample(0:999, 250, replace=T) 
```

Supposex = (x1,x2,...,xn) denotes the vector xVec andy = (y1,y2,...,yn) denotes the vector yVec.

(a) Create the vector (y2 −x1,...,yn −xn−1). Plot the vector.

```{r}
yVeca <- yVec[-1]#delete the first element
xVeca <- xVec[-length(xVec)]#dropping the last one, can also use"-250" instead of "-length"
dVec <- yVeca - xVeca
plot(dVec, col="red")
```

(b) Create the vector (sin(y1)/cos(x2), sin(y2)/cos(x3), sin(yn-1)/cos(xn))

```{r}
yVecb <- yVec[-length(yVec)]
xVecb <- xVec[-1]
eVec <- sin(yVecb)/cos(xVecb)
eVec
```

(c) Create the vector (x1 +2x2 −x3, x2 +2x3 −x4, ... , xn−2 +2xn−1 −xn). 

```{r}
xVecc1 <- xVec[-c((length(xVec)-1):length(xVec))]
xVecc2 <- xVec[-c(1, length(xVec))]
xVecc3 <- xVec[-c(1, 2)]
fVec <- xVecc1 + 2*xVecc2 - xVecc3
fVec
```

(d) Calculate summary of (e^(-x(i-1)))/(xi+10) as i goes from 1 to n-1.

```{r}
e <- 2.718281828459
xVecd1 <- xVec[-1]
xVecd2 <- xVec[-length(xVec)]
gVec <- (e^(-xVecd1))/(xVecd2+10)
gVec
```


7. This question uses the vectors "xVec"" and "yVec"" created in the previous question and the functions "sort"", "order", "mean", "sqrt", "sum" and "abs".

```{r}
set.seed(50)
xVec <- sample(0:999, 250, replace=T)
yVec <- sample(0:999, 250, replace=T)
```

(a) Pick out the values in yVec which are > 600. Make a chart that shows the data points with values > 600 as a subset.

```{r}
plot(yVec, col="red")
bVec <- yVec[yVec>600]
points(bVec, col="green")
bVec
```

(b) What are the index positions in yVec of the values which are > 600?

```{r}
i <- (1:250)
c <- (601:999)
n <- c(is.element(yVec[i], c)*1)
n
```

(c) What are the values in xVec which correspond to the values in yVec which are > 600? (By correspond, we mean at the same index positions.) 

```{r}
m <- xVec*n
m
```

(d) Create the vector (sqrt(|x1 - xbar|), sqrt(|x2 - xbar|), ... , sqrt(|xn - xbar|)) where xbar denotes the mean of the vector x = (x1, x2, ... , xn). Make a chart that shows the xVec and the vector you created.

```{r}
i <- (1:250)
xbar <- mean(xVec[i])
xbar
v <- c(sqrt(abs(xVec[i] - xbar)))
v
```

(e) How many values in yVec are within 200 of the maximum value of the terms in yVec?

```{r}
i <- (1:250)
q <- max(yVec[i]) - 200
q
c <- (q:max(yVec[i]))
n <- c(is.element(yVec[i], c)*1)
sum(n[i])
```

(f) How many numbers in xVec are divisible by 2? (Note that the modulo operator is denoted %%.)

```{r}
i <- (1:250)
z <- c(is.element(xVec[i]%%2,0)*1)
sum(z[i])
```

(g) Sort the numbers in the vector xVec in the order of increasing values in yVec. Make a chart.

```{r}
g <- sort(xVec, decreasing=FALSE)
g
```

(h) Pick out the elements in yVec at index positions 1,4,7,10,13,.... Make a chart.

```{r}
i <- seq(1, 250, by=3)
h <- c(yVec[i])
h
```

8. By using the function "cumprod" or otherwise, calculate 1+2/3+2/3*4/5+2/3*4/5*6/7+ ... +2/3*4/5* ... *38/39

```{r}
i <- c(1:19)
a <- cumprod(2*i)
b <- cumprod(2*i+1)
j <- 1+sum(a[i]/b[i])
j
```


## Exercise 2 Matrices


1.(a)

```{r}
A <- matrix(data = c(1,1,3,5,2,6,-2,-1,-3), nrow = 3, byrow = TRUE) 
A
```

```{r}
A%*%A%*%A
```

(b)

```{r}
Ab <- A
Ab[7:9] <- Ab[4:6]+A[7:9]
Ab
```


2.

```{r}
b <- c(10, -10, 10)
B <- matrix(data = c(rep(b, time=15)), nrow = 15, byrow = TRUE)
B
```

```{r}
BT <- t(B) #Transfer B
BT
```

```{r}
BT %*% B
```


3.

```{r}
matE <- matrix(data = seq(0, 0, length=36), nrow = 6, byrow = TRUE)
matE
row(matE)
```

```{r}
c <- c(0, 1, 0, 0, 0, 0, 1)
C <- matrix(data=rep(c, len=36), nrow = 6, byrow = TRUE)
C
#or
#A <- matrix(0, nrow=6, ncol=6, byrow=TRUE)
#(abs(col(A)-row(A))==1)*1
```


4.

```{r}
outer(0:4,0:4,"+") #if no "+" then default is multiplication
```


5.
(a)

```{r}
outer(0:4,0:4,"+") %% 5
```

(b)

```{r}
outer(0:9,0:9,"+") %% 10
```

(c)

```{r}
outer(0:8,9:1,"+") %% 9
```

6.

```{r}
A <- abs(outer(0:4,1:5,"-") + 1) + 1
b <- c(7, -1, -3, 5, 17)
# solve Ax = b
x <- solve(A,b)
x
```


7.

```{r}
set.seed(75)
aMat <- matrix( sample(10, size=60, replace=T), nr=6)
aMat
```

(a)

```{r}
c <- (5:10)
i <- is.element(aMat, c) * 1
aMatb <- matrix(data = i, nrow = 6, byrow = TRUE)
aMatb
```

```{r}
n1 <- sum(aMatb[1,1:10])
n2 <- sum(aMatb[2,1:10])
n3 <- sum(aMatb[3,1:10])
n4 <- sum(aMatb[4,1:10])
n5 <- sum(aMatb[5,1:10])
n6 <- sum(aMatb[6,1:10])
n <- c(n1, n2, n3, n4, n5, n6)
n
```

(b)

```{r}
i <- is.element(aMat, 7) * 1
aMatc <- matrix(data = i, nrow = 6, byrow = TRUE)
aMatc
```

```{r}
n1 <- sum(aMatc[1,1:10])
n2 <- sum(aMatc[2,1:10])
n3 <- sum(aMatc[3,1:10])
n4 <- sum(aMatc[4,1:10])
n5 <- sum(aMatc[5,1:10])
n6 <- sum(aMatc[6,1:10])
n <- c(n1, n2, n3, n4, n5, n6)
n
```

```{r}
is.element(n, 2)*1
```

(c)

```{r}

```


8.

(a)

```{r}
i <- c(1:20)
j <- c(1:5)
sum((i^4)/(3+j))
```

(b)

```{r}
i <- c(1:20)
j <- c(1:5)
sum((i^4)/(3+i*j))
```

(c)

```{r}
sum <- 0
for (i in 1:20) {
  for (j in 1:i) {
    sum <- sum + (i^4)/(3+i*j)
  }
}
sum
```

