---
title: "Problem Set 3"
author: "Munan Hou"
date: "November 8,  2016"
output: html_document
---

```{r}
library(dplyr)
hipp_data <- read.csv("C:/YUKIHO/BU/MA 615/Classes/9.12.2016/class 9.12.2016/ma_568/Homework 3/hipp_data.csv")
```

#Problem 1

```{r}
plot(hipp_data$xN, hipp_data$yN, type='l', xlab="coordinate x", ylab="coordinate y")

hipp1_data <- subset(hipp_data, spikes==1)
points(hipp1_data$xN, hipp1_data$yN, col='red')

hipp2_data <- subset(hipp_data, spikes2==1)
points(hipp2_data$xN, hipp2_data$yN, col='blue')
```

The plot shows that, both neuron's spikes are gathered in the left-bottom of the graph; thus we could say they are probably correlated.

#Problem 2

```{r}
m1 <- glm(spikes ~ xN  +  yN, poisson, hipp_data)
summary(m1)
```

All parameters are significant.

```{r}
lambda <- function(x, y) {
  exp(m1$coefficients[1] + m1$coefficients[2] * x + m1$coefficients[3] * y)
  }

x2 <- y2 <- seq(-1, 1, .1)
?outer
z<- outer(x2, y2, lambda)
persp(x2,  y2,  z,  theta = 30,  phi = 30,  expand = 0.5,  col = "lightblue",  xlab="X",  ylab = "Y", main="Maximum Likelihood Model Indensity") -> res
phi <- seq(0,  2 * pi,  len = 201)
xr <- cos(phi)
yr <- sin(phi)
lines(trans3d(xr, yr, lambda(xr, yr), res),  col = "pink",  lwd = 2)
```

Due to the graph shows, the spatial firing properties are well captured.

#Problem 3

```{r}
hipp_data$xN2 <- (hipp_data$xN)^2
hipp_data$yN2 <- (hipp_data$yN)^2
m2 <- glm(spikes ~ xN2 + yN2 + xN + yN, poisson, hipp_data)
summary(m2)
```

All significant.

```{r}
lambda2 <- function(x, y) {
  exp(m2$coefficients[1] + m2$coefficients[2] * (x^2) + m2$coefficients[3] * (y^2) +  m2$coefficients[4] * x + m2$coefficients[5] * y)
  }

x2 <- y2 <- seq(-1, 1, .1)
z2<- outer(x2, y2, lambda2)
persp(x2,  y2,  z2,  theta = 30,  phi = 30,  expand = 0.5,  col = "lightblue",  xlab="X",  ylab = "Y", main="Maximum Likelihood Model Indensity") -> res
phi <- seq(0,  2 * pi,  len = 201)
xr <- cos(phi)
yr <- sin(phi)
lines(trans3d(xr, yr, lambda(xr, yr), res),  col = "pink",  lwd = 2)
```

```{r}
m1$aic
m2$aic
```

#Problem 4

```{r}
hipp_data_n <- hipp_data[, c("spikes", "xN", "yN", "xN2", "yN2", "spikes_hist1", "spikes_hist2", "spikes_hist3", "spikes_hist4", "spikes_hist5", "spikes_hist6", "spikes_hist7", "spikes_hist8", "spikes_hist9", "spikes_hist10", "spikes_hist11", "spikes_hist12", "spikes_hist13", "spikes_hist14", "spikes_hist15", "spikes_hist16", "spikes_hist17", "spikes_hist18", "spikes_hist19", "spikes_hist20")]
p <- aic <- c()
for (i in 6:25) {
  hipp_data_a <- hipp_data_n[, 1:i]
  m3 <- glm(spikes ~ ., poisson, hipp_data_a)
  aic <- c(aic, m3$aic)
  p <- c(p, i-5)
  }

rec <- data.frame("p"=p, "aic"=aic)
plot(rec$p, rec$aic, type="l")
points(rec$p, rec$aic)
```

```{r}
subset(rec,  aic==min(aic))
```

#Problem 5

```{r}
hipp_data_n2 <- hipp_data[, c("spikes", "xN", "yN", "xN2", "yN2", "spikes_hist1", "spikes_hist2", "spikes_hist3", "spikes_hist4", "spikes_hist5", "spikes_hist6", "spikes_hist7", "spikes_hist8", "spikes2_hist1", "spikes2_hist2", "spikes2_hist3", "spikes2_hist4", "spikes2_hist5", "spikes2_hist6", "spikes2_hist7", "spikes2_hist8", "spikes2_hist9", "spikes2_hist10", "spikes2_hist11", "spikes2_hist12", "spikes2_hist13", "spikes2_hist14", "spikes2_hist15", "spikes2_hist16", "spikes2_hist17", "spikes2_hist18", "spikes2_hist19", "spikes2_hist20")]

p2 <- aic2 <- c()

for (i in 14:33) {
  hipp_data_a <- hipp_data_n2[, 1:i]
  m3 <- glm(spikes ~ ., poisson, hipp_data_a)
  aic2 <- c(aic2, m3$aic)
  p2 <- c(p2, i-13)
  }

rec2 <- data.frame("p1"=8, "p2"=p2, "aic2"=aic2)

subset(rec2,  aic2==min(aic2))
```

#Problem 6

```{r}
hipp_data_n3 <- select(hipp_data, spikes, xN, yN, xN2, yN2, spikes_hist1:spikes_hist20,  spikes2_hist1:spikes2_hist20)
p41 <- p42 <- aic4 <- c()

for (k in 6:25){
  hipp_data_a <- hipp_data_n3[, 1:k] %>% select(-xN, -yN, -xN2, -yN2)
  m3 <- glm(spikes ~ ., poisson, hipp_data_a)
  aic4 <- c(aic4, m3$aic)
  p41 <- c(p41, k-5)
  p42 <- c(p42, 0)
  for (i in 26:45){
    hipp_data_a <- hipp_data_n3[, c(1:k, 26:i)] %>% select(-xN, -yN, -xN2, -yN2)
    m3 <- glm(spikes ~ ., poisson, hipp_data_a)
    aic4 <- c(aic4, m3$aic)
    p41 <- c(p41, k-5)
    p42 <- c(p42, i-25)
    }
  }

rec4 <- data.frame("p1"=p41, "p2"=p42, "aic"=aic4)
subset(rec4, aic==min(aic))
```

This does have a minimum aic. Almost every parameters are significant. The reason for why only when we fail to model the spatial component of the firing activiety is that, in spatial component, the correlation between history and network interaction is more significant.

#Problem 7

```{r}
m0 <- glm(spikes ~ xN2 + yN2 + xN + yN + spikes_hist1 + spikes_hist2 + spikes_hist3 + spikes_hist4  + spikes_hist5 + spikes_hist6 + spikes_hist7 + spikes_hist8,  poisson, hipp_data)
hipp_data[, "num"] <- 1:nrow(hipp_data)
hipp_data_zz <- subset(hipp_data, spikes==1)
Xi <- hipp_data_zz[, "num"]

```
